---
title: "Course Project No. 4"
author: "Cuachon, Guzman"
date: "May 25, 2019"
output: html_document
---

This is our first Markdown document.

Project No. 4 for Stat 197 (Data Science)

# Setting Directory
```{r echo=TRUE}
setwd("C:/Users/CLIENT/Desktop")

```

#Reading the datasets

```{r dataset, echo=TRUE}
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
```

#Exploring the PM2.5 Emissions Data


```{r NEI, echo=TRUE}
head(NEI)
names(NEI)
```

#Exploring the Source Classification Code Table Data

```{r SCC, echo=TRUE}
head(SCC)
names(SCC)
```


#Questions

##Question 1: Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

###Codes and plot for Question 1
```{r plot1.R, echo=TRUE}
total <- aggregate(Emissions ~ year, NEI, sum)
barplot(
    (total$Emissions)/10^6,
    names.arg=total$year,
    xlab="Year",
    ylab="PM2.5 Emissions (10^6 Tons)",
    main="Total PM2.5 Emissions From All US Sources"
)
```


##Question 2: Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (fips=="24510") from 1999 to 2008? Use the base plotting system to make a plot answering this question.

###Codes and plot for Question 2
```{r plot2.R, echo=TRUE}
baltimoreNEI <- NEI[NEI$fips=="24510",]
aggTotalsBaltimore <- aggregate(Emissions ~ year, baltimoreNEI,sum)
barplot(
    aggTotalsBaltimore$Emissions,
    names.arg=aggTotalsBaltimore$year,
    xlab="Year",
    ylab="PM2.5 Emissions (Tons)",
    main="Total PM2.5 Emissions From All Baltimore City Sources"
)
```


##Question 3: Of the four types of sources indicated by the type (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999-2008 for Baltimore City? Which have seen increases in emissions from 1999-2008? Use the ggplot2 plotting system to make a plot answer this question.

###Codes and plot for Question 3
```{r plot3.R, echo=TRUE}
#Aggregating and naming the columns
balem <-aggregate(NEI$Emissions,by=list(NEI$year, NEI$fips=='24510',NEI$type),sum)
names(balem) <-c('year','Bal','type','Emissions')
balem<-balem[balem$Bal==TRUE,]

#Plotting
library(ggplot2)
qplot(year,Emissions,data=balem, geom=c('line','point'), color=type, 
      main='Emissions from 1999-2008 for Baltimore City by Type', ylab='Emissions (tons)')
```

##Question 4: Across the United States, how have emissions from coal combustion-related sources changed from 1999-2008?

###Codes and plot for Question 4
```{r plot4.R, echo=TRUE}
#Getting the SCC for coal-related combustions
i<-grep('Coal',SCC$EI.Sector,ignore.case = TRUE)
index <-SCC[i,]$SCC

#Aggregating and naming the columns
coalem <-aggregate(NEI$Emissions,by=list(NEI$year, NEI$SCC %in% index),sum)
names(coalem) <-c('year','Coal','Emissions')
coal.em<-coalem[coalem$Coal==TRUE,]

#Plotting
library(ggplot2)
qplot(year,Emissions,data=coal.em, geom=c('line','point'), 
      main='Emission from Coal Combustion-Related Sources 1999-2008', ylab='Emissions (tons)')
```

##Question 5: How have emissions from motor vehicle sources changed from 1999-2008 in Baltimore City?

###Codes and plot for Question 5
```{r plot5.R, echo=TRUE}


motorem <-aggregate(NEI$Emissions,
                    by=list(NEI$year, NEI$fips=='24510', NEI$type=='ON-ROAD'),sum)
names(motorem) <-c('year','type','Bal','Emissions')
motor.em<-motorem[motorem$type==TRUE & motorem$Bal==TRUE,]

#Plotting
library(ggplot2)
qplot(year,Emissions,data=motor.em, geom=c('line','point'), 
      main='Emission from Motor-Related Sources 1999-2008 in Baltimore City',
      ylab='Emissions (tons)')
```


##Question 6: Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (fips == "06037").Which city has seen greater changes over time in motor vehicle emmissions?

###Codes and plot for Question 6
```{r plot6.R, echo=TRUE}
motor <-aggregate(NEI$Emissions,
                    by=list(NEI$year, NEI$type=='ON-ROAD',NEI$fips),sum)
names(motor) <-c('year','type','fips','Emissions')
motorem.city<-motor[motor$type==TRUE & 
                     (motor$fips=='24510'|motor$fips=='06037'),]
motorem.city$city <-ifelse(motorem.city$fips=='24510','Baltimore City','LA County')


#Create changes variable with the baseline being Emissions in 1999 for both cities

baseline<-motorem.city[motorem.city$year==1999,]
motorem.city$baseline<-ifelse(motorem.city$fips=='24510',baseline[baseline$fips=='24510',4],
                         baseline[baseline$fips=='06037',4])
motorem.city$changes<-motorem.city$Emissions/motorem.city$baseline

#Plot
library(ggplot2)
qplot(year,changes,data=motorem.city, geom=c('line','point'), 
      main='Changes in Emission from Motor-Related Sources 1999-2008 
      Baltimore City vs LA County',
      color=city, ylab='Changes in Emissions (Emissions in 1999 =1)')
```



